### 2024-11-16 Article/Background_sections/Adhesivity.tex

\begin{cor}\label{cor:adhesivity_functor_categories}
    Let $\cat A$ be a $\mathcal M$-adhesive category for some $\mathcal M \subseteq \Hom(\cat A)$. Then, for every other category $\cat C$, the functor category $[\cat C, \cat A]$ is $\mathcal M^{\cat C}$-adhesive, where \todo{Devi chiedere che C abbia tutti i pullback}
    \[
        \mathcal{M}^\cat C = \{\eta \in \Hom([\cat C, \cat A]) \mid \eta_C \in \mathcal M \textit{ for each object $C$ of $\cat C$ }\}
    \]
\end{cor}

>>>>

\begin{cor}\label{cor:adhesivity_functor_categories}
    Let $\cat A$ be a $\mathcal M$-adhesive category for some $\mathcal M \subseteq \Hom(\cat A)$. Then, if $\cat C$ is a category whith all the pullbacks, the functor category $[\cat C, \cat A]$ is $\mathcal M^{\cat C}$-adhesive, where
    \[
        \mathcal{M}^\cat C = \{\eta \in \Hom([\cat C, \cat A]) \mid \eta_C \in \mathcal M \textit{ for each object $C$ of $\cat C$ }\}
    \]
\end{cor}

### 2024-11-16 Article/Background_sections/Basic_Notions.tex
minori
eqgrph pushouts reg mono van kampen

%    Let now be $f = f_i \circ f_s$ be the decomposition of $f$ with $f_s$ surjective (i.e., epi in $\Set$) and $f_i$ injective (i.e., mono in $\Set$), and $g = g_i \circ g_s$ be the decomposition of $g$ with $g_i$ injective and $g_s$ surjective, having the following situation.
%    \[
%        \begin{tikzcd}
%            X \ar[r, "{f_s}"] \ar[d, "e"] & f(X) \ar[r, "{f_i}"] & M \ar[d, "m"] \\
%            Y \ar[r, "{g_i}"swap] \ar[urr, "h"]& Y/_\sim \ar[r, "{g_s}"swap] & Z
%        \end{tikzcd}
%   \]
%    For the diagram above to commute, must be $f_i \circ f_s = h \circ e$ and $g_s \circ g_i = m \circ h$\todo{completare!!!}
%\end{proof}}
\begin{proof}
	Let $y \in Y$. Since $e$ is epi (i.e., surjective), there exists $x \in X$ such that $e(x) = y$. Then, we put $h(y) = f(x)$. $h$ is well defined, in fact, let $x' \in X$ be such that $e(x') = y$. In this case, we have $h(x) = f(x')$, and so
	\begin{align*}
		(m \circ f) (x')  &= (g \circ e) (x') \\
				&= g(y)		\\
				&= (g \circ e) (x)	\\
				&= (m \circ f) (x)
	\end{align*}
	since $m$ is mono by hypothesis, $f(x) = f(x')$.
\end{proof}

\subsection{Categories from other categories}\label{ssect:cats_from_cats}

>>>>

%    Let now be $f = f_i \circ f_s$ be the decomposition of $f$ with $f_s$ surjective (i.e., epi in $\Set$) and $f_i$ injective (i.e., mono in $\Set$), and $g = g_i \circ g_s$ be the decomposition of $g$ with $g_i$ injective and $g_s$ surjective, having the following situation.
%    \[
%        \begin{tikzcd}
%            X \ar[r, "{f_s}"] \ar[d, "e"] & f(X) \ar[r, "{f_i}"] & M \ar[d, "m"] \\
%            Y \ar[r, "{g_i}"swap] \ar[urr, "h"]& Y/_\sim \ar[r, "{g_s}"swap] & Z
%        \end{tikzcd}
%   \]
%    For the diagram above to commute, must be $f_i \circ f_s = h \circ e$ and $g_s \circ g_i = m \circ h$\todo{completare!!!}
%\end{proof}}
\begin{proof}
	Let $y \in Y$. Since $e$ is epi (i.e., surjective), there exists $x \in X$ such that $e(x) = y$. Then, we put $h(y) = f(x)$. $h$ is well defined, in fact, let $x' \in X$ be such that $e(x') = y$. In this case, we have $h(x) = f(x')$, and so
	\begin{align*}
		(m \circ f) (x')  &= (g \circ e) (x') \\
				&= g(y)		\\
				&= (g \circ e) (x)	\\
				&= (m \circ f) (x)
	\end{align*}
	since $m$ is mono by hypothesis, $f(x) = f(x')$.
\end{proof}
{\color{red}
\begin{cor}\label{cor:unique}
	% La fattorizzazione di un morfismo è unica a meno di isomorfismo \todo{Scrivilo bene (l'isomorfismo dovrà rispettare certe coerenze) e dimostralo}
	In $\Set$, let $f:A \to B$ be a morphism, and let $f=m \circ e$, $f=m'\circ e'$ be two factorizations of $f$, with $e: A \to X$ and $e': A \to Y$ epis and $m: X \to B$, $m': Y \to B$ monos.
	Then, there exists an isomorphism $\phi: X \to Y$.
\end{cor}

\begin{proof}
	Representing the hypothesis in diagrams, we have the following situation
	\[\begin{tikzcd}
		A \ar[r, "e"] \ar[d, "e'"swap] & X \ar[d, "m"] & A \ar[r, "e'"] \ar[d, "e"swap] & Y \ar[d, "m'"] \\
		Y \ar[r, "m'"swap] \ar[ur, dashed, "h"] & B    & X \ar[r, "m"swap] \ar[ur, dashed, "k"] & B 
	\end{tikzcd}\]
	where the dashed arrows are yielded by \Cref{prop:prop_epi_mono_Set}. Hence, we have that
	\[\begin{split}
		id_X \circ e &= h \circ e' \\ &= h \circ k \circ e
	\end{split}\qquad\begin{split}
		id_Y \circ e' &= k \circ e \\ &= k \circ h \circ e'
	\end{split}\]
	Hence, $h \circ k = id_X$, $k \circ h = id_Y$, obtaining the isomorphism.
\end{proof}
}
\subsection{Categories from other categories}\label{ssect:cats_from_cats}

### 2024-11-16 Article/Graph_sections/Eq_Graphs_integration.tex
proof of 2.2.12

Another functor that will be useful later is $V: \EqGrph \to \Set$, sending $(E_G, V_G, C_G, s_G, t_G, q_G)$ to $C_G$, and $h = (h_E, h_V, h_C)$ to $h_C$.

@@

	For $C'_1$, we can make a similar argument, let $S$ be the pullback of $m_2$ along $x_1$, using \Cref{lemma:pullback_lemma} and, again, the proof of \Cref{prop:eqgrph_complete} we know that $q'_1$ arise as the factorization of the arrow $B'_1\to S$ induced by $q'_2\circ h'_2$ and $a_2$ so that we have a diagram.
	\[\begin{tikzcd}[row sep = 20, column sep =20]
	{B'_1} && {B'_2} \\
	\\
	{C'_1} && S && {C'_2} \\
	\\
	{B_1} && T && U \\
	\\
	&& {C_1} && {C_2}
	\arrow["{h'_V}", from=1-1, to=1-3]
	\arrow["{q'_1}", from=1-1, to=3-1]
	\arrow["{q'_2}", from=1-3, to=3-5, bend left = 20]
	\arrow["{m_1}", from=3-1, to=3-3]
	\arrow["{s_1}", from=3-3, to=3-5]
	\arrow["{s_2}"', from=3-3, to=5-3]
	\arrow["{m_2}", from=3-5, to=5-5]
	\arrow["{q_1}"', from=5-1, to=7-3, bend right = 20, shift right = 1]
	\arrow["{x_1}", from=5-3, to=5-5]
	\arrow["{x_2}"', from=5-3, to=7-3]
	\arrow["{u_2}", from=5-5, to=7-5]
	\arrow["{h_C}"', from=7-3, to=7-5]
        \arrow["{a_V}"', from=1-1, to=5-1, bend right = 20, shift right = 1]
	\end{tikzcd}\]

	Moreover, we have that
	\[
		\begin{split}
			s_1 \circ m_1 \circ q_1 &= q'_2 \circ h'_V \\ &= h'_C \circ q'_1
		\end{split}\qquad\begin{split}
			q'_4 \circ t'_V &= t'_C \circ q'_2\\ &= u_1\circ m_2 \circ q'_2
		\end{split}
	\]\[
		\begin{split}
			y_1 \circ m_3 \circ q'_3 &= p'_V \circ q'_4\\ &= p'_C \circ q'_3
		\end{split}\qquad\begin{split}TODO:\ w \circ s_2 \circ m_1 = m_3 \circ k_C'
		\end{split}
	\]
	Hence, $s_1\circ m_1 = h'_C$, $w \circ s_2 \circ m_1 = m_3 \circ k_C'$, $t'_C = u_1 \circ m_2$ and $p'_C = y_1 \circ m_3$.
	Let now $z: \graph{G}'_2 \to \graph H = (E, V, Q, s, t, q)$, $w: \graph G_3^1 \to \graph H$ be two morphisms such that $z \circ h' = w \circ k'$, and let $\phi:B'_4 \to V$ be the arrow induced by $z_V$ and $w_V$. We want to construct the arrow $v: C'_4 \to Q$ in the diagram below.
	\[\begin{tikzcd}[row sep = 25, column sep = 25]
	& {B_1'} && {B_2'} \\
	{B_3'} && {B_4'} && V \\
	& T && U \\
	Y && {C_4'} && Q \\
	& {C_1} && {C_2} \\
	{C_3} && {C_4}
	\arrow["{h_V'}", from=1-2, to=1-4]
	\arrow["{k_V'}"', from=1-2, to=2-1]
	\arrow["{s_2\circ m_1 \circ q_1'}"{description, pos=0.7}, from=1-2, to=3-2]
	\arrow["{t_V'}", from=1-4, to=2-3]
	\arrow["{z_V}", from=1-4, to=2-5]
	\arrow["{m_2\circ q'_2}"{pos=0.7}, from=1-4, to=3-4]
	\arrow["{m_3 \circ q'_3}"', from=2-1, to=4-1]
	\arrow["q", from=2-5, to=4-5]
	\arrow["{x_1}"{pos=0.8}, from=3-2, to=3-4]
	\arrow["w"', from=3-2, to=4-1]
	\arrow["{x_2}"'{pos=0.7}, from=3-2, to=5-2]
	\arrow["{u_1}", from=3-4, to=4-3]
	\arrow["{u_2}"{pos=0.7}, from=3-4, to=5-4]
	\arrow["{y_2}"', from=4-1, to=6-1]
	\arrow["{h_C}"{pos=0.8}, from=5-2, to=5-4]
	\arrow["{k_C}", from=5-2, to=6-1]
	\arrow["{t_C}", from=5-4, to=6-3]
	\arrow["{p_C}"', from=6-1, to=6-3]
	\arrow["{p_V'}"{pos=0.8}, from=2-1, to=2-3, crossing over]
	\arrow["\phi"{pos=0.8}, from=2-3, to=2-5, crossing over]
	\arrow["{q'_4}"'{pos=0.7}, from=2-3, to=4-3, crossing over]
	\arrow["{y_1}"{pos=0.7}, from=4-1, to=4-3, crossing over]
	\arrow["{d_C}"'{pos=0.7}, from=4-3, to=6-3, crossing over]
	\arrow["v"{pos=0.8}, dashed, from=4-3, to=4-5, crossing over]
	\end{tikzcd}\]

>>>>

Another functor that will be useful later is $V: \EqGrph \to \Set$, sending $(E_G, V_G, C_G, s_G, t_G, q_G)$ to $C_G$ and $h = (h_E, h_V, h_C)$ to $h_C$.

@@

	For $C'_1$, we can make a similar argument, let $S$ be the pullback of $m_2$ along $x_1$, using \Cref{lemma:pullback_lemma} and, again, the proof of \Cref{prop:eqgrph_complete} we know that $q'_1$ arise as the factorization of the arrow $B'_1\to S$ induced by $q'_2\circ h'_2$ and $a_2$ so that we have a diagram.
	\[\begin{tikzcd}[row sep = 16, column sep =16]
	{B'_1} && {B'_2} \\
	\\
	{C'_1} && S && {C'_2} \\
	\\
	{B_1} && T && U \\
	\\
	&& {C_1} && {C_2}
	\arrow["{h'_V}", from=1-1, to=1-3]
	\arrow["{q'_1}", from=1-1, to=3-1]
	\arrow["{q'_2}", from=1-3, to=3-5, bend left = 20]
	\arrow["{m_1}", from=3-1, to=3-3]
	\arrow["{s_1}", from=3-3, to=3-5]
	\arrow["{s_2}"', from=3-3, to=5-3]
	\arrow["{m_2}", from=3-5, to=5-5]
	\arrow["{q_1}"', from=5-1, to=7-3, bend right = 20, shift right = 1]
	\arrow["{x_1}", from=5-3, to=5-5]
	\arrow["{x_2}"', from=5-3, to=7-3]
	\arrow["{u_2}", from=5-5, to=7-5]
	\arrow["{h_C}"', from=7-3, to=7-5]
        \arrow["{a_V}"', from=1-1, to=5-1, bend right = 20, shift right = 1]
	\end{tikzcd}\]

	Moreover, we have that
	\[
		\begin{split}
			s_1 \circ m_1 \circ q_1 &= q'_2 \circ h'_V \\ &= h'_C \circ q'_1
		\end{split}\qquad\begin{split}
			q'_4 \circ t'_V &= t'_C \circ q'_2\\ &= u_1\circ m_2 \circ q'_2
		\end{split}
	\]\[
		\begin{split}
			y_1 \circ m_3 \circ q'_3 &= p'_V \circ q'_4\\ &= p'_C \circ q'_3
		\end{split}\qquad\begin{split}TODO:\ w \circ s_2 \circ m_1 = m_3 \circ k_C'
		\end{split}
	\]
	Hence, $s_1\circ m_1 = h'_C$, $w \circ s_2 \circ m_1 = m_3 \circ k_C'$, $t'_C = u_1 \circ m_2$ and $p'_C = y_1 \circ m_3$.
	Let now $z: \graph{G}'_2 \to \graph H = (E, V, Q, s, t, q)$, $w: \graph G_3^1 \to \graph H$ be two morphisms such that $z \circ h' = w \circ k'$, and let $\phi:B'_4 \to V$ be the arrow induced by $z_V$ and $w_V$. We want to construct the arrow $v: C'_4 \to Q$ in the diagram below.
	\[\begin{tikzcd}[row sep = 20, column sep = 20]
	& {B_1'} && {B_2'} \\
	{B_3'} && {B_4'} && V \\
	& T && U \\
	Y && {C_4'} && Q \\
	& {C_1} && {C_2} \\
	{C_3} && {C_4}
	\arrow["{h_V'}", from=1-2, to=1-4]
	\arrow["{k_V'}"', from=1-2, to=2-1]
	\arrow["{s_2\circ m_1 \circ q_1'}"{description, pos=0.7}, from=1-2, to=3-2]
	\arrow["{t_V'}", from=1-4, to=2-3]
	\arrow["{z_V}", from=1-4, to=2-5]
	\arrow["{m_2\circ q'_2}"{pos=0.7}, from=1-4, to=3-4]
	\arrow["{m_3 \circ q'_3}"', from=2-1, to=4-1]
	\arrow["q", from=2-5, to=4-5]
	\arrow["{x_1}"{pos=0.8}, from=3-2, to=3-4]
	\arrow["w"', from=3-2, to=4-1]
	\arrow["{x_2}"'{pos=0.7}, from=3-2, to=5-2]
	\arrow["{u_1}", from=3-4, to=4-3]
	\arrow["{u_2}"{pos=0.7}, from=3-4, to=5-4]
	\arrow["{y_2}"', from=4-1, to=6-1]
	\arrow["{h_C}"{pos=0.8}, from=5-2, to=5-4]
	\arrow["{k_C}", from=5-2, to=6-1]
	\arrow["{t_C}", from=5-4, to=6-3]
	\arrow["{p_C}"', from=6-1, to=6-3]
	\arrow["{p_V'}"{pos=0.8}, from=2-1, to=2-3, crossing over]
	\arrow["\phi"{pos=0.8}, from=2-3, to=2-5, crossing over]
	\arrow["{q'_4}"'{pos=0.7}, from=2-3, to=4-3, crossing over]
	\arrow["{y_1}"{pos=0.7}, from=4-1, to=4-3, crossing over]
	\arrow["{d_C}"'{pos=0.7}, from=4-3, to=6-3, crossing over]
	\arrow["v"{pos=0.8}, dashed, from=4-3, to=4-5, crossing over]
	\end{tikzcd}\]
